Feature Selection and Classification of \\ Spam on Social Networking Sites

### 2012-05-09 docs/paper/paper.tex
update

We enjoyed unrestricted access to the data of InterPals.net, a SNS with 
over 1.3 million active members. This data includes a corpus of over a 
100 million private messages and another 2 million messages that have been 
labeled as spam by users. Other data includes 40 million or so ``wall" 
comments, 5 million photos, and 8 million photo comments.

@@

In the data processing step, we extracted the domain name from each email address. We found that ham user email domains 
were distributed across top email providers in similar proportions (Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
predominance of Yahoo email accounts among spammers can be seen in Figure \ref{fig:email}. While we have no certain 
explanation for this, Figure \ref{fig:yspam} shows that Ghana and Nigeria alone account for 38\% of the IP 
addresses associated with these Yahoo accounts. This popularity of Yahoo among West African users has 
been noted before\cite{burrell}, and appears to be due to the early penetration of Yahoo services in the region.

@@

In order to train our classifiers, we created a feature matrix from the message and account data. 
We did this using Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document 
frequencies in descending order, mapping each feature (word or expert feature value) to an integer 
key. The second step was to create a sparse matrix representation of each document (in this case, 
spam or ham message) and the associated features. Spark provides a parallelized in-memory computation 
framework that dramatically reduced the time necessary to extract these features, build the dictionary 
and generate the feature matrix.

@@

The Naive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
the features that we used, \eg{account lifetime, and whether a user has friends or photos}, are in 
fact likely to be co-dependent, leading to an overweight of these features during classification.

>>>>

We enjoyed unrestricted access to the data of InterPals.net, a SNS with 
over 1.3 million active members. The data from this site includes a corpus of over a 
100 million private messages and another 2 million messages that have been 
labeled as spam  and deleted by users. Other data includes 40 million or so ``wall" 
comments, 5 million photos and 8 million photo comments.

@@

In the data processing step, we extracted the domain name from each email address. We found that ham user email domains 
were distributed across top email providers in similar proportions (Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
predominance of Yahoo email accounts among spammers can be seen in Figure \ref{fig:email}. While we have no certain 
explanation for this, Figure \ref{fig:yspam} shows that Ghana and Nigeria alone account for 38\% of the IP 
addresses associated with these Yahoo accounts. This popularity of Yahoo among West African users has 
been noted before\cite{burrell} and appears to be due to the early penetration of Yahoo services in the region.

@@

In order to train our classifiers, we created a feature matrix from the message and account data. 
We did this using Perl, Spark and Lucene analyzers. The first step was to generate a dictionary sorted by document 
frequencies in descending order, mapping each feature (word or expert feature value) to an integer 
key. The second step was to create a sparse matrix representation of each document (in this case, 
spam or ham message) and the associated features. Spark provides a parallelized in-memory computation 
framework that dramatically reduced the time necessary to extract these features, build the dictionary 
and generate the feature matrix.

@@

The Naive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
the features that we used, \eg{account lifetime and whether a user has friends or photos}, are in 
fact likely to be co-dependent, leading to an overweight of these features during classification.
